---
title: "MSRS LME with Random Effects"
author: "Harshi Gupta"
date: "12/14/2019"
output:
  html_document: default
  pdf_document: default
---

## Setup

In this section, please change the "SC_category" variable to the correct concentration. R script is run 5 times to get all the concentrations.
```{r Setup, warning = FALSE, tidy=TRUE, message = FALSE}
###################################################
### Set Working Directory
setwd("C:/Users/HGupta/Documents/Hopkins/MSRS/2020/Data/Results_Copy")

### clear the workspace
rm(list=ls())

###################################################
### Load libraries
library(lme4)
library(lattice)
library(reshape2)
library(dplyr)
library(openxlsx)
library(readxl)
library(stats)
library(data.table)
library(psych)
library(ggplot2)
library(knitr)

# Ensure PDF isn't cut off
opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)

# File naming information
SC_category = "clinical"

# Import the names of the judges for the given SC_category (i.e. concentration) above.
raw_judges <- read.table(file = paste(SC_category,"_judges.txt", sep = ""),header = FALSE, sep="\t", quote="",stringsAsFactors = FALSE)
# Check to see if this looks right.
raw_judges
```


Import the raw data, which includes each judges' scores for each abstract. 
```{r, warning = FALSE, tidy=TRUE, message = FALSE}
# ########### Data
# ## Import the raw data from the Excel file. Check Github repository for proper format for input file.
all_data <- read_excel("all_raw.xlsx")
summary(all_data)
# Reorder columns. Add the total score for judges and remove extraneous columns.
all_data$Score <- rowSums(all_data[3:6])
all_data <- all_data[-c(3:6)]

# Rename columns in dataframe to Judge, ID, Score
colnames(all_data) <- c("Judge","ID","Score")

# Extract values for judges in category assigned above
raw_data <- all_data[all_data$Judge %in% raw_judges[,1],]

# Order dataframe by name of Judge
raw_data <- raw_data[order(raw_data$Judge),]


adjusted_data <- raw_data #rename raw data

# Convert ID column to a character class.
adjusted_data$ID <- as.character(adjusted_data$ID)

str(adjusted_data)
```

## Linear Mixed Model

The following code will do following things:  
1) Run LME with random effects.  
Dependent variable: Score  
Fixed effect: ID (Abstract)  
Random effect: Judges  
2) Add column for LMER-Adjusted Score and Z-score (based on ID and Judge)  
3) Average the Adjusted Scores for each ID and generate a 2-column list with ID, LMER-Adjusted, and Z-Scores
```{r LME, warning = FALSE, tidy=TRUE, message = FALSE}
# We have a data-frame adjusted_data with variables Judge, ID, and Score 
#    ID has values of 3-digit ID numbers
#    Judge has values by names
#    Score has values from 4 (all 1's) to 28 (all 7's)

##### Fit a linear-mixed-model with random effects for the judges

# LMER model generated
lmer_model <- lmer(Score ~ factor(ID) + (1 | Judge), data = adjusted_data)
#lme_model2 <-lme(behaviour ~ task*sex, random = ~ 1|ID/task, method="ML", data=dat)

summary(lmer_model)

# Add adjusted LMER scores corresponding to each actual score as a column, removing the random effect.
adjusted_data$Score_LMER <- predict(lmer_model, re.form=NA, data = adjusted_data)
```


```{r Graphs, warning = FALSE, tidy=TRUE, message = FALSE}
###### Make caterpillar plot of conditional SD of random variables (Judge)
randoms <- ranef(lmer_model, postVar = TRUE)
qq <- attr(ranef(lmer_model, postVar = TRUE)[[1]], "postVar")
rand.interc <- randoms$Judge
df <- data.frame(Intercepts = randoms$Judge[, 1], sd.interc = 2 * sqrt(qq[,, 1:length(qq)]), lev.names = rownames(rand.interc))
df$lev.names <- factor(df$lev.names, levels = df$lev.names[order(df$Intercepts)])
p <- ggplot(df, aes(lev.names, Intercepts, shape = lev.names))

# Added horizontal line at y=0, error bars to points and points with size two
p <- p + geom_hline(yintercept = 0) + geom_errorbar(aes(ymin = Intercepts - sd.interc, ymax = Intercepts + sd.interc), width = 0, color = "black") + geom_point(aes(size = 2))

# Removed legends and with scale_shape_manual point shapes set to 1 and 16
p <- p + guides(size = FALSE, shape = FALSE) + scale_shape_manual(values = c(rep(x = 1,nrow(df[df$Intercepts < 0,])), rep(x = 16, nrow(df[df$Intercepts > 0,]))))

#Changed appearance of plot (black and white theme) and x and y axis labels
p <- p + theme_bw() + xlab("Levels:Judges") + ylab("Random Effects")

#Final adjustments of plot
p <- p + theme(axis.text.x=element_text(size=rel(1.2)),
               axis.title.x=element_text(size=rel(1.3)),
               axis.text.y=element_text(size=rel(1.2)),
               panel.grid.minor=element_blank(),
               panel.grid.major.x=element_blank())

#To put levels on y axis you need to use coord_flip()
p <- p+ coord_flip()
print(p)

###### Make residual plot
plot(lmer_model)
```


```{r Final, warning = FALSE, tidy=TRUE, message = FALSE}
##### Add column of z-scores for each student based on judge average and SD
adjusted_data <- setDT(adjusted_data)[, Z.score := scale(Score), Judge]

### Create a file that has the adjusted scores and raw scores.
write.xlsx(adjusted_data, file = paste(SC_category,"_","raw_adjusted.xlsx", sep = ""), append = FALSE, sep = "\t", dec = ".", row.names = FALSE, col.names = TRUE, quote = FALSE)

### Make a ranked table grouped by IDs and corresponding average adjusted and raw scores
RANKED_DATA <- adjusted_data %>% group_by(ID) %>%
                        summarise(Avg_Score_LMER = mean(Score_LMER),Avg_Z_score = mean(Z.score), Avg_raw_score = mean(Score)) %>%
                        ungroup() %>%
                        arrange(desc(Avg_Score_LMER)) %>%
                        as.data.frame()

write.xlsx(RANKED_DATA, file = paste(SC_category,"_","means_adjusted.xlsx", sep = ""), append = FALSE, sep = "\t", dec = ".", row.names = FALSE, col.names = TRUE, quote = FALSE)
```
